---
title: "regression"
author: "Elisabeth Nesmith"
date: "12/13/2021"
output: html_document
---

```{r}
library(tidyverse)

# load in hms_data_new
hms_data <- read_csv("Data/hms_data_new.csv")
```

```{r}
# transform response variables to binary variables
hms_data <- hms_data %>% 
  mutate(aca_impa_bin = ifelse(aca_impa == 1 | aca_impa == 2, 0, 1)) %>% 
  mutate(belong_bin = ifelse(belong1 == 1 | belong1 == 2 | belong1 == 3, 0, 1)) %>% 
  mutate(discrim_bin = ifelse(discrim == 1, 0, 1))

# convert race NAs to 0s
hms_data[, 12:19][is.na(hms_data[, 12:19])] <- 0

# convert variables to factors
factors <- c(4:25)

hms_data[, 4:25] <- lapply(hms_data[,factors], factor)
```

# Modeling Academic Impact

```{r}
# run model for academic impact, controlling for race and fincur, trans is predictor variable
aca_impa_model <- glm(data = hms_data, formula = aca_impa_bin ~ race_black + race_ainaan + race_asian + race_his_temp + race_pi + race_mides + race_white + race_other + fincur + trans)

summary(aca_impa_model)
```

```{r}
# odds ratio
exp(coef(aca_impa_model))
```

A trans student is 1.2 times more likely to have experienced 3 or more academic impact days than a cis student.

# Modeling Sense of Belonging
```{r}
# run model for sense of belonging, controlling for race and fincur, trans is predictor variable
belong_model <- glm(data = hms_data, formula = belong_bin ~ race_black + race_ainaan + race_asian + race_his_temp + race_pi + race_mides + race_white + race_other + fincur + trans)

summary(belong_model)
```

```{r}
# odds ratio
exp(coef(belong_model))
```

A trans student is 1.09 times more likely to feel as if they do not belong on campus than a cis student.

# Modeling Discrimination

```{r}
# run model for discrimination, controlling for race and fincur, trans is predictor variable
discrim_model <- glm(data = hms_data, formula = discrim_bin ~ race_black + race_ainaan + race_asian + race_his_temp + race_pi + race_mides + race_white + race_other + fincur + trans)

summary(discrim_model)
```

```{r}
# odds ratio
exp(coef(discrim_model))
```

A trans student is 1.3 times more likely to experience discrimination than a cis student.